%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
   Progress is possible only if we train ourselves to think about programs\\
   without thinking of them as pieces of executable code.
} \\ \medskip
    % - E. Dijkstra [[l12][¹²]]
    %--- \defcitealias{knuth:1974}{Donald E. Knuth}\citetalias{knuth:1974} \citep{knuth:1974}
    --- \defcitealias{dijkstra2000pursuit}{Edsgar W. Dijkstra}\citetalias{dijkstra2000pursuit} \citep{dijkstra2000pursuit}
\end{flushright}



\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
I thank my advisor Dr. Piyush P. Kurur for introducing me to the world
of programming language theory (along with emacs), and without whom
this thesis would not exist.
I am also thankful for the guidance that I received from
Dr. Sandeep Chandran especially for getting familiar with hardware
aspects needed for this work.
I thank the teachers, staff and students of IIT Palakkad, especially
those from the Department of Computer Science and Engineering for
their support.
I am thankful to all the people that I have interacted with online and
offline, especially Max, who helped me better understand how to navigate
research life.
I thank my family for all their support and encouragement during my MS.
% During the course of my MS, I have benefited from conversations with a
% wide variety of people both online and offline. espec
I am also grateful to the communities of Coq and Clash for their help
and suggestions while I was struggling with errors.

\endgroup
